---
title: "R Objects"
author: 'Lutao DAI'
date: "Aug 20, 2019"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
  html_document:
    code_folding: show
    highlight: haddock
    theme: readthedown #yeti #spacelab #simplex #readable #paper #flatly #cosmo #lumen
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---

```{r, include=F}
require(pacman)
pacman::p_load(tidyverse)
```


## Comparison
Comparison operators:  >, >=, <, <=, == (equal), != (not equal)

```{r}
sqrt(4)^2 == 4
```

```{r}
sqrt(2)^2 == 2
```
```{r}
near(sqrt(2)^2, 2)
```


## Logical Operators
$ (and), | (or), ! (not), xor

## Missing Values
Missing values are extremely problematic. Almost any operations involving `NA` will result in `NA`.
```{r}
NA + 3
```
```{r}
na.value <- NA
na.value == NA
```

Use `is.na` to check if a value is missing
```{r}
is.na(na.value)
```

Generally, there are three ways of treating them.
1. Get rid of them completely. For example, remove columns that contain too many missing values, or drop rows that with missing values
1. Preserve observations or columns but create indicators
1. (Use with care) Replace missing values with a numerical value as placeholder. For example, since value -1 is meaningless for the column `AGE`, you can replace `NA` in the column `AGE` with -1. However, be careful if you try to perform numerical operations on them.